### 2014-08-03 tex/GAME_HARD_01_CONCLUSION.tex
Corrected spelling throughout document.

In conclusion; this study proposes, pursuant to the established performance of WARP, that WARP-like technologies are feasable for use where the reference~device~driver is too slow, when hardware is not available, and for extended purposes other than debugging and profiling.

% CONCLUSION - FUTURE WORK
\subsection{Future~Work}
\label{sec:conclusion:futurework}
The performance improvements of WARP, in comparison to the reference~device~driver, calls for further inquiry into what the flaws of using WARP may be.
Microsoft lists, amongst other limitations latency-oriented CPUs may have over throughput oriented GPU architectures, memory bandwidth as a potential bottleneck for WARP-type acceleration methods~\citeweb[]{warp}.
In accordance to the suspicions raised by Microsoft, it may be benificial to investigate the performance of WARP when such bottlenecks are stressed.
Furthermore, the author suggests complementary elaboration into double precision calculations, with the intent of examining whether or not the Subjects detailed in section~\nameref{sec:contribution:subjectsofstudy} may have differentiating effects on computational precision.

>>>>

In conclusion; this study proposes, pursuant to the established performance of WARP, that WARP-like technologies are feasible for use where the reference~device~driver is too slow, when hardware is not available, and for extended purposes other than debugging and profiling.

% CONCLUSION - FUTURE WORK
\subsection{Future~Work}
\label{sec:conclusion:futurework}
The performance improvements of WARP, in comparison to the reference~device~driver, calls for further inquiry into what the flaws of using WARP may be.
Microsoft lists, amongst other limitations latency-oriented CPUs may have over throughput oriented GPU architectures, memory bandwidth as a potential bottleneck for WARP-type acceleration methods~\citeweb[]{warp}.
In accordance to the suspicions raised by Microsoft, it may be beneficial to investigate the performance of WARP when such bottlenecks are stressed.
Furthermore, the author suggests complementary elaboration into double precision calculations, with the intent of examining whether or not the Subjects detailed in section~\nameref{sec:contribution:subjectsofstudy} may have differentiating effects on computational precision.

### 2014-08-03 tex/GAME_HARD_01_CONTRIBUTION.tex
Corrected spelling throughout document.

\subsubsection{Windows Advanced Rasterization Platform (CPU)}
The simulated execution of a DirectCompute kernel using a specially devised software rasterizer devoloped for high performance and full framework conformance by Microsoft in the latest version of DirectX.
Being based on the reference~device~driver, WARP utilizes thread pooling to distribute tasks efficiently on the CPU, along with grouping execution in batches for optimum performance.
Furthermore, WARP uses JIT~(Just-In-Time) compilation to generate SSE2 and SSE4.1 SIMD~(Single Instruction, Multiple Data) instructions, in addition to x86-compatible native instructions to significantly speed up simulation of GPU-bound workloads~\citeweb[]{warp}.
Having outperformed some low-end integrated GPUs according to Microsoft, and the device driver being included in Windows~7 and 8 runtimes, Microsoft recommends using the driver for retail applications, such as casual games~\citeweb[]{warp}.
Although Microsoft has presented some performance comparisons to hardware acceleration, there is no performance benchmark indicative of WARP performance in relation to the reference~device~driver - the device commonly used for debugging and verification purposes.
As such, expectations on the performance of WARP is unclear, but the driver is expected to perform better than the reference~device~driver from which it is derived.\\

@@

% CONTRIBUTION - KERNELS
\subsection{Kernels}
\label{sec:contribution:kernels}
In addition to the Subject drivers mentioned in section~\nameref{sec:contribution:subjectsofstudy}, two kernels with varying level of optimization are examined.
The kernels are implemented in accordance to the CUDA matrix multiplication kernels as described by Kirk~and~Hwu~\cite[p.~67, p.~87]{Kirk:2010:PMP:1841511}.
Furthermore, aforementioned kernels accomodate both integer- and floating point precision.
Descriptions of these kernels is presented below, and are complemented with source code under section~\nameref{sec:appendix}.

@@

\subsubsection{Matrix mult. w. Thread Blocks and Shared Memory}
Similar to the previously described kernel, but optimized to utilize shared memory in order to reduce time-consuming access of global memory, as presented by Kirk~and~Hwu~\cite[p.~77-93]{Kirk:2010:PMP:1841511}.
Stratton~et~al.~\cite[p.~1-3]{Stratton:2008:MEI:1485701.1485703} instructs that the CUDA GPGPU-model may be applied onto multicore CPUs, including locality-wise execution of logical thread blocks (that is; all threads in a block limited to a single core), with the utilization of local- and shared-memory approximately corresponding to the L1-cache of a CPU core.
Hence, the kernel is presented as a scenario due to the preconditions of WARP, stating that a kernel optimized for GPU-execution is likewise optimized for execution with WARP~\citeweb[]{warp}.
Thus, we investigate a more optimized kernel to see whether or not this behaviour may be replicated in the experiment in terms of memory access.
The program will be referred to as the Tiled~Kernel throughout the material.

@@

\subsubsection{analytics}
A utility devised to compose data surrounding possible precisional deviation between matrices $C$~and~$Ref$.
The tool, written in C\texttt{++}, compiles the minimum- and maximum  deviation encountered, as well as to calculate the standard deviation of any precisional discrepency in produced product matrices.
In turn, analytics outputs this information to an intermediate file.\\

@@

% CONTRIBUTION - PROCESS OF STUDY
\subsection{Process of Study}
\label{sec:contribution:processofstudy}
For the purpose of this study, the Object~of~Study, being $200\times 200$ matrices of integer or floating point precision, is randomized with numbers between zero and ten.
The product matrix of these matrices is subsequently computed $100$ times for each configuration.
In this way, the Basic~Kernel is executed with each Subject~of~Study - being Hardware~Acceleration, the reference~device~driver, and WARP - respectively, as is the Tiled~Kernel.
For each execution, data surrounding the dispatch time of each kernel (regardless of program initialization) is garnered along with precision-wise deviational data.
This process is repeated for integer- and floating point precision, in order to establish if there is any performance-wise disrepency betwixt them.

@@

In terms of memory latency, table~\ref{tab:contribution:results:summaryint} demonstrates performance gains with the hardware accelerated Subject (when utilizing shared memory amongst blocks), with varying results for the other Subjects; either increasing or decreasing execution time (see table~\ref{tab:contribution:results:summaryfloat}).
As such, applicable to both integer- and floating point precision, the performance of WARP is impaired by the kernel utilizing shared memory.
In the Microsoft guide on WARP~\citeweb[]{warp}, the author claims that an application tuned to run efficiently on hardware will run efficiently whilst simulated using WARP as well - and vice versa.
However, the collected data rather indicate a roughly $20\%$ percentual increase in execution time, independant of precision, even though the same kernel accelerates the hardware accelerated Subject by roughly $70$\%.
This demonstrates that that the capabilities of WARP do not extend to memory latency optimization by lower-latency shared memory.
The floating point scenario of this effect is visualized in figure~\ref{fig:contribution:results:warp:msswarp}.

Analysis indicates no loss in precision between reference- and product matrices.
Neither is there any recorded divergence in results compiled by respective Subjects.
In hindsight, this behaviour may be expected as WARP conforms to the precision requirements of the Direct3D~$10$- and $10.1$-specification~\citeweb[]{warp}.
See the Microsoft guide on Floating-point Rules~\citeweb[]{floatingpointrules} for more information surrounding floating point precision in the Direct3D-framework.
Accordingly, albeit not being explicitly developed for the purposes of precision, WARP performs equally as with the reference~device~driver.
Whether or not this behaviour is replicated for double floating point precision is uncertain.

\begin{figure}[htb]
\begin{center}
	\resizebox{ \columnwidth }{!}{\input{../msswarp}}
	\caption{Execution time using WARP with floating point-precision for the Basic and Tiled kernels, indicated by their respective mean- and seperated by standard deviation values. Samples outside of respective standard deviation are not presented.}
	\label{fig:contribution:results:warp:msswarp}
\end{center}
\end{figure}

>>>>

\subsubsection{Windows Advanced Rasterization Platform (CPU)}
The simulated execution of a DirectCompute kernel using a specially devised software rasterizer developed for high performance and full framework conformance by Microsoft in the latest version of DirectX.
Being based on the reference~device~driver, WARP utilizes thread pooling to distribute tasks efficiently on the CPU, along with grouping execution in batches for optimum performance.
Furthermore, WARP uses JIT~(Just-In-Time) compilation to generate SSE2 and SSE4.1 SIMD~(Single Instruction, Multiple Data) instructions, in addition to x86-compatible native instructions to significantly speed up simulation of GPU-bound workloads~\citeweb[]{warp}.
Having outperformed some low-end integrated GPUs according to Microsoft, and the device driver being included in Windows~7 and 8 run-times, Microsoft recommends using the driver for retail applications, such as casual games~\citeweb[]{warp}.
Although Microsoft has presented some performance comparisons to hardware acceleration, there is no performance benchmark indicative of WARP performance in relation to the reference~device~driver - the device commonly used for debugging and verification purposes.
As such, expectations on the performance of WARP is unclear, but the driver is expected to perform better than the reference~device~driver from which it is derived.\\

@@

% CONTRIBUTION - KERNELS
\subsection{Kernels}
\label{sec:contribution:kernels}
In addition to the Subject drivers mentioned in section~\nameref{sec:contribution:subjectsofstudy}, two kernels with varying level of optimization are examined.
The kernels are implemented in accordance to the CUDA matrix multiplication kernels as described by Kirk~and~Hwu~\cite[p.~67, p.~87]{Kirk:2010:PMP:1841511}.
Furthermore, aforementioned kernels accommodate both integer- and floating point precision.
Descriptions of these kernels is presented below, and are complemented with source code under section~\nameref{sec:appendix}.

@@

\subsubsection{Matrix mult. w. Thread Blocks and Shared Memory}
Similar to the previously described kernel, but optimized to utilize shared memory in order to reduce time-consuming access of global memory, as presented by Kirk~and~Hwu~\cite[p.~77-93]{Kirk:2010:PMP:1841511}.
Stratton~et~al.~\cite[p.~1-3]{Stratton:2008:MEI:1485701.1485703} instructs that the CUDA GPGPU-model may be applied onto multicore CPUs, including locality-wise execution of logical thread blocks (that is; all threads in a block limited to a single core), with the utilization of local- and shared-memory approximately corresponding to the L1-cache of a CPU core.
Hence, the kernel is presented as a scenario due to the preconditions of WARP, stating that a kernel optimized for GPU-execution is likewise optimized for execution with WARP~\citeweb[]{warp}.
Thus, we investigate a more optimized kernel to see whether or not this behavior may be replicated in the experiment in terms of memory access.
The program will be referred to as the Tiled~Kernel throughout the material.

@@

\subsubsection{analytics}
A utility devised to compose data surrounding possible precisional deviation between matrices $C$~and~$Ref$.
The tool, written in C\texttt{++}, compiles the minimum- and maximum  deviation encountered, as well as to calculate the standard deviation of any precisional discrepancy in produced product matrices.
In turn, analytics outputs this information to an intermediate file.\\

@@

% CONTRIBUTION - PROCESS OF STUDY
\subsection{Process of Study}
\label{sec:contribution:processofstudy}
For the purpose of this study, the Object~of~Study, being $200\times 200$ matrices of integer or floating point precision, is randomized with numbers between zero and ten.
The product matrix of these matrices is subsequently computed $100$ times for each configuration.
In this way, the Basic~Kernel is executed with each Subject~of~Study - being Hardware~Acceleration, the reference~device~driver, and WARP - respectively, as is the Tiled~Kernel.
For each execution, data surrounding the dispatch time of each kernel (regardless of program initialization) is garnered along with precision-wise deviation data.
This process is repeated for integer- and floating point precision, in order to establish if there is any performance-wise discrepancy betwixt them.

@@

In terms of memory latency, table~\ref{tab:contribution:results:summaryint} demonstrates performance gains with the hardware accelerated Subject (when utilizing shared memory amongst blocks), with varying results for the other Subjects; either increasing or decreasing execution time (see table~\ref{tab:contribution:results:summaryfloat}).
As such, applicable to both integer- and floating point precision, the performance of WARP is impaired by the kernel utilizing shared memory.
In the Microsoft guide on WARP~\citeweb[]{warp}, the author claims that an application tuned to run efficiently on hardware will run efficiently whilst simulated using WARP as well - and vice versa.
However, the collected data rather indicate a roughly $20\%$ percentual increase in execution time, independent of precision, even though the same kernel accelerates the hardware accelerated Subject by roughly $70$\%.
This demonstrates that that the capabilities of WARP do not extend to memory latency optimization by lower-latency shared memory.
The floating point scenario of this effect is visualized in figure~\ref{fig:contribution:results:warp:msswarp}.

Analysis indicates no loss in precision between reference- and product matrices.
Neither is there any recorded divergence in results compiled by respective Subjects.
In hindsight, this behavior may be expected as WARP conforms to the precision requirements of the Direct3D~$10$- and $10.1$-specification~\citeweb[]{warp}.
See the Microsoft guide on Floating-point Rules~\citeweb[]{floatingpointrules} for more information surrounding floating point precision in the Direct3D-framework.
Accordingly, albeit not being explicitly developed for the purposes of precision, WARP performs equally as with the reference~device~driver.
Whether or not this behavior is replicated for double floating point precision is uncertain.

\begin{figure}[htb]
\begin{center}
	\resizebox{ \columnwidth }{!}{\input{../msswarp}}
	\caption{Execution time using WARP with floating point-precision for the Basic and Tiled kernels, indicated by their respective mean- and separated by standard deviation values. Samples outside of respective standard deviation are not presented.}
	\label{fig:contribution:results:warp:msswarp}
\end{center}
\end{figure}
